Consistent Polyhedral Surrogates via Embedding

### 2019-05-21 notes/jessie-proof-sketches.tex
Another iteration of Jessie's proofs

  \jessie{Redone.  Original version in comments below.}
  For the forward implication, suppose $L$ embeds the discrete loss $\ell$ via the embedding $\varphi$.
  Then $\Gamma := \prop{L}$ embeds the finite property $\gamma:\simplex \toto \R$ elicited by $\ell$.
  Call $\R' := \varphi(\R)$.
  %For all $p \in \simplex$, we know that $\varphi(\gamma(p)) \subseteq \Gamma(p)$ \raf{Not true; focus on the next statement and show directly using def of embedding}.
  We know $\gamma' := \Gamma(p) \cap \R'$ is nonempty for all $p \in \simplex$ since $\varphi(\gamma(p)) \subseteq \R' \subseteq \Gamma(p)$.
  By \jessie{COLT Lemma 11}, we know that $L|_{\R'}$ elicits $\gamma'$ and $\risk{L} = \risk{L|_{\R'}}$.

@@

  For the reverse implication, assume that $\risk{L} = \risk{\ell}$.
  Since $\ell$ is discrete, we know these risks are polyhedral, as $\risk{\ell}$ is the pointwise min of a finite set of linear functions.
  The projection of their epigraphs onto $\simplex$ must form the same power diagram \jessie{COLT Theorem 33}.  %, corresponding to the level sets of the properties each loss elicits.
  Observe there are only finitely many cells in a power diagram; we call the set of full-dimensional cells in the power diagram $F$.
  We claim that each cell $f \in F$ is then a level set $\gamma_r$ for some $r \in \R$, and equivalently, $\Gamma_{u_f}$ for some $u_f \in \reals^d$.
  We go on to embed $\varphi: r \mapsto u_f$ for each $r \in \R$.

  Now, for each $f \in F$, fix some distribution $p \in \inter{f}$.
  The supporting hyperplane of $\risk{\ell}(p)$ is unique, since $p$ belongs to exactly one facet of $\risk{\ell}$ (i.e. $\risk{\ell}(p)$ is differentiable.)
  Since $\ell$ is discrete, the coefficients of the supporting hyperplane exactly describe the loss over each outcome given the report $r = \gamma(p)$.  \jessie{??}
  %Similarly, for some $u \in \reals^d$, as the risks are the same, we claim that $L(u,p) = \ell(\gamma(p), p)$, as the risks are equal.

  As there is some $u_f \in \argmin_{u'} L(u', p)$, we can design $\varphi : r \mapsto u_f \in \Gamma(p)$, and by the risks being equal, we also have $r \in \argmin_{r' \in \R} \ell(r', p)$, and thus $r \in \gamma(p)$.
  The reverse implication follows similarly, and we conclude that $\Gamma$ embeds $\gamma$.

  Since $\risk{L}$ is polyhedral, the supporting hyperplane to $\risk{L}$ is unique for each $p \in \inter{f}$ and $f \in F$, and so is the loss $L(u_f)$ is unique for all such $p$.
  As both the loss values of $\ell(r)$ and $L(u_f)$ are equal and minimize their respective losses,  so we conclude that $L$ embeds $\ell$.

@@

\begin{proposition}\label{prop:embed-risk-poly}
  A loss $L$ embeds a discrete loss if and only if $\risk{L}$ is polyhedral.
  \jessie{Ready for a look over.}
\end{proposition}
\begin{proof}[Property version]
  For the forward direction, suppose $L$ embeds a discrete loss $\ell$.
  By Proposition~\ref{prop:embed-bayes-risks}, we know that $\risk{L} = \risk{\ell}$.
  Since $\ell$ is discrete, we know that $\risk{\ell}$ is polyhedral as $\risk{\ell}$ is the max of a finite set of linear functions in $\simplex$. \raf{do we know? okay, we do---but why?}.
  Therefore, $\risk{L}$ is polyhedral.
%  Suppose $\Gamma := \prop{L}$ embeds the finite property $\gamma$, elicited by the discrete loss $\ell$.
%  We know that $\risk{\ell}$ is polyhedral as $\gamma$ is finite.
%  Since $L$ embeds $\ell$, we know $\risk{L} = \risk{\ell}$, and therefore $\risk{L}$ is also polyhedral. 
%  \jessie{This direction is shorter below without properties.}

  Now suppose $\risk{L}$ is polyhedral, and $\Gamma := \prop{L}$.
  Consider, as above, the power diagram induced by projecting $\risk{L}$ onto $\simplex$ corresponding to the level sets of $\Gamma$.
  Call the set of full-dimensional cells in this power diagram $\F$, and note that $\F$ is finite since it defined by a projection of a polyhedral function.
  We know that $\trim(\Gamma)$ is a finite set, and its full dimensional level sets union to the simplex.
  We can then invoke \jessie{COLT Prop 12} to observe that $\Gamma$ embeds a finite property.
  In particular, when the embedding is the identity, we know that the loss values match up for each $r \in \R$, so we can conclude that $L$  embeds a discrete loss.
  \raf{I would have thought we could use COLT Prop 12 at this point in the proof... take a look at its proof, the 2 $\Rightarrow$ 3 part, and note that it already invokes Lemma 11, so it already proves this direction}
  %For each $f \in \F$, choose some $p \in \inter{f}$ and add some $u_f \in \argmin_u L(u, p)$ to the set $\R$.
  %As level sets are closed, we know that $\Gamma(p) \cap \R$ is nonempty for all $p \in \simplex$ \raf{how does this follow from closure?}.
  %By \jessie{COLT Lemma 11}, we know $\Gamma$ embeds $\Gamma \cap \R$ elicited by the discrete loss $L|_\R$. 
  %Thus, $L$ embeds a discrete loss-- namely, $L|_\R$.
\end{proof}
%\hrulefill
%\begin{proof}[Non-property version]
%  For the forward direction, suppose $L$ embeds a discrete loss $\ell$.
%  By Proposition~\ref{prop:embed-bayes-risks}, we know that $\risk{L} = \risk{\ell}$.
%  Since $\ell$ is discrete, we know that $\risk{\ell}$ is polyhedral.
%  Therefore, $\risk{L}$ is polyhedral.
%
%  For the reverse implication, suppose $\risk{L}$ is polyhedral.
%  For each facet $f \in F$, consider the report $u_f$ as constructed above.
%  Since $F$ is finite, the set $\R = \{u_f\}_{f \in F}$ is finite.
%  Now define the discrete loss $\ell:\R \to \Y$ by $\ell(r,y) = L(r,y)$.
%
%  To see that $L$ embeds this discrete $\ell$, consider the embedding function to be the identity for all $r \in \R$.
%  By Proposition~\ref{prop:embed-bayes-risks}, it is sufficient to show $\risk{L} = \risk{\ell}$.
%  For all $p \in \simplex$, we observe
%  \begin{align*}
%  	\risk{L}(p) = \inf_{u \in \reals^d} L(u,p) = \min_{r \in \R} L(r,p) = \min_{r \in \R} \ell(r, p) = \risk{\ell}(p)
%  \end{align*}
%  Note that the second equality holds because for every $p \in \simplex$, we know there is some $r \in \R$ such that $r \in \arginf_u L(u,p)$ by closure of subgradient sets, and thus the loss values are the same, regardless of if we restrict to $\R$ or stay in $\reals^d$.
%  Thus, $L$ embeds the discrete $\ell$ as constructed by Proposition~\ref{prop:embed-bayes-risks}.
%
%  \raf{Just take the discrete reports to be the facets, and use the Bayes risk result.}
%\end{proof}
%\hrule 
\hrulefill

@@

\begin{theorem}\label{thm:poly-embeds-discrete}
  Every polyhedral loss $L$ embeds a discrete loss.
  \jessie{My thoughts are in a comment, and I think Raf said it was wrong at one point, but I don't understand why exactly.}
\end{theorem}
\begin{proof}
  We merely show $L$ polyhedral $\implies$ $\risk{L}$ polyhedral, and apply Proposition~\ref{prop:embed-risk-poly}.
  \raf{Not sure how trivial this is, but it's definitely true.  I suspect it has a very short proof.}
  \jessie{$L(u,p)$ polyhedral in $\reals^d$ for all $p \in \simplex$ implies $L^*(0,p) = -\risk{L}(p)$ polyhedral in $\simplex$ implies $\risk{L}$ polyhedral.
  Now we can apply Proposition~\ref{prop:embed-risk-poly} to observe the result.}

>>>>

  \jessie{Redone.  Original version in comments below.}
  For the forward implication, suppose $L$ embeds the discrete loss $\ell$ via the embedding $\varphi$.
  Then $\Gamma := \prop{L}$ embeds the finite property $\gamma:\simplex \toto \R$ elicited by $\ell$.
  Call $\U := \varphi(\R)$.
  %For all $p \in \simplex$, we know that $\varphi(\gamma(p)) \subseteq \Gamma(p)$ \raf{Not true; focus on the next statement and show directly using def of embedding}.
  $\gamma'(p) := \Gamma(p) \cap \U$ is nonempty for all $p \in \simplex$ because $r \in \gamma(p) \implies \varphi(r) \in \Gamma(p)$ for all $p \in \simplex$, and by definition $varphi(r) \in \U$, so $\varphi(r) \in \gamma'(p)$ for all $p \in \simplex$.
  \jessie{Possible rephrase: $\gamma'(p) := \Gamma(p) \cap \U$ is nonempty for all $p \in \simplex$ as $\varphi(r) \in \gamma'(p)$ by the definition of embedding and nondegeneracy of $\gamma$.}
  By \jessie{COLT Lemma 11}, we know that $L|_{\U}$ elicits $\gamma'$ and $\risk{L} = \risk{L|_{\U}}$.

@@

  For the reverse implication, assume that $\risk{L} = \risk{\ell}$.
  Since $\ell$ is discrete, we know these risks are polyhedral, as $\risk{\ell}$ is the pointwise min of a finite set of linear functions.
  The projection of their epigraphs onto $\simplex$ forms a power diagram \jessie{COLT Theorem 33}, and since the risks are the same, we know the corresponding power diagrams are equal.  %, corresponding to the level sets of the properties each loss elicits.
  Observe there are only finitely many cells in a power diagram; we call the set of full-dimensional cells in the power diagram $F$.
  We claim that each cell $f \in F$ is then a level set $\gamma_r$ for some $r \in \R$, and equivalently, $\Gamma_{u_f}$ for some $u_f \in \reals^d$.
  We go on to embed $\varphi: r \mapsto u_f$ for each $r \in \R$.
  (By regularity assumptions, we consider that there is an $f \in F$ corresponding to each $r \in \R$.)

  Now, for each $f \in F$, fix some distribution $p \in \inter{f}$.
  The supporting hyperplane of $\risk{\ell}(p)$ is unique, since $p$ belongs to exactly one facet of $\risk{\ell}$ (i.e. $\risk{\ell}(p)$ is differentiable.)
  Since $\ell$ is discrete, the coefficients of the supporting hyperplane (in $\simplex \subseteq \reals^n$) exactly describe the loss over each outcome given the report $r \in \gamma(p)$.  \jessie{Double check??}

  Now choose some $u_f \in \argmin_{u'} L(u', p)$; we design $\varphi : r \mapsto u_f \in \Gamma(p)$.
  Since $\risk{L}$ is polyhedral, the supporting hyperplane to $\risk{L}$ is unique for each $p \in \inter{f}$ and $f \in F$, and so the loss $L(u_f)$ is equal to $\risk{L}(p)$ for all such $p$.
  \jessie{Don't think I phrased this correctly. revisit.}
  Therefore, $u_f \in \Gamma(p)$ for all $p \in f$ since level sets of a property are closed.

  We claim that $L$ embeds $\ell$, since we have $r \in \argmin_{r' \in \R} \ell(r', p) \iff \varphi(r) \in \argmin_{u' \in \U} L(u', p)$ by our design of $\varphi$.
  Moreover, the equlity of $\ell(r)$ and $L(\varphi(r))$ can be seen by the risks being the same, and therefore so are the supporting hyperplanes (defining the loss values at $\ell(r)$ and $L(\varphi(r))$) for all $p \in f$ and $f \in F$, since $\cup_F f = \simplex$.

@@

\begin{proposition}\label{prop:embed-risk-poly}
  A loss $L$ embeds a discrete loss if and only if $\risk{L}$ is polyhedral.
  \jessie{Ready for a look over.}
\end{proposition}
\begin{proof}
  For the forward direction, suppose $L$ embeds a discrete loss $\ell$.
  By Proposition~\ref{prop:embed-bayes-risks}, we know that $\risk{L} = \risk{\ell}$.
  Since $\ell$ is discrete, we know that $\risk{\ell}$ is polyhedral as $\risk{\ell}$ is the max of a finite set of linear functions in $\simplex$. \raf{do we know? okay, we do---but why?}.
  Therefore, $\risk{L}$ is polyhedral.

  The reverse direction follows similarly to the proof of Proposition~\ref{prop:embed-bayes-risks}.
  Consider, as above, the power diagram induced by projecting $\risk{L}$ onto $\simplex$ corresponding to the level sets of $\Gamma$.
  Call the set of full-dimensional cells in this power diagram $F$, and note that $F$ is finite since it defined by a projection of a polyhedral function.
  For each $f \in F$, there is a distribution $p \in \inter{f}$ such that the supporting hyperplane at $\risk{L}(p)$ is unique.
  Pick some $u_f \in \argmin_{u' \in \U}L(u', p)$.
  The coefficients of the supporting hyperplane of $\risk{L}(p)$ exactly describe the discrete loss $\ell(u_f)$.
  Repeat this process for every facet, and denote $\R := \{u_f\}_{f \in F}$.
  Since $\cup f = \simplex$, and each $f = \Gamma_{u_f}$ for some $u_f \in \reals^d$, we invoke Proposition~\ref{prop:embed-trim} \jessie{COLT Prop 12} to note $\Gamma:= \prop{L}$ embeds a finite property $\gamma := \prop{ell}$, and their loss values match up, since $\ell$ is exactly described by the same supporting hyperplanes for each $p \in \simplex$.

%  Now suppose $\risk{L}$ is polyhedral, and $\Gamma := \prop{L}$.
%  Consider, as above, the power diagram induced by projecting $\risk{L}$ onto $\simplex$ corresponding to the level sets of $\Gamma$.
%  Call the set of full-dimensional cells in this power diagram $\F$, and note that $\F$ is finite since it defined by a projection of a polyhedral function.
%  We know that $\trim(\Gamma)$ is a finite set, and its full dimensional level sets union to the simplex.
%  We can then invoke \jessie{COLT Prop 12} to observe that $\Gamma$ embeds a finite property.
%  In particular, when the embedding is the identity, we know that the loss values match up for each $r \in \R$, so we can conclude that $L$  embeds a discrete loss.
%  \raf{I would have thought we could use COLT Prop 12 at this point in the proof... take a look at its proof, the 2 $\Rightarrow$ 3 part, and note that it already invokes Lemma 11, so it already proves this direction}

\end{proof}
\hrulefill

@@

\begin{theorem}\label{thm:poly-embeds-discrete}
  Every polyhedral loss $L$ embeds a discrete loss.
  \jessie{My thoughts are in a comment, and I think Raf said it was wrong at one point, but I don't understand why exactly.}
\end{theorem}
\begin{proof}
  We merely show $L$ polyhedral $\implies$ $\risk{L}$ polyhedral, and apply Proposition~\ref{prop:embed-risk-poly}.
  \raf{Not sure how trivial this is, but it's definitely true.  I suspect it has a very short proof.}
  \jessie{Follows as a Corollary of Lemma 3 in NeurIPS paper.}
